---
title: "Assignment3"
author: "RF"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}

library(pacman)
p_load(stringr, dplyr, ggplot2, gridExtra, brms, devtools, mvtnorm, rethinking, Brobdingnag)
setwd("C:/Users/Bruger/Desktop/Cognitive Science/R/Portfolio-3_part1")
d = read.csv("Assignment3Data.csv")



?subset
typic = subset(d, ASD ==0)
autist = subset(d, ASD ==1)

#visit 1
v1 = subset(d, Visit ==1)


```

## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

The questions you will have to answer are:

1. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results. P.S. Should you scale? P.P.S. Don't forget to motivate your priors. P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.
1.1. Verbal IQ and ADOS
1.2. Non Verbal IQ and ADOS
1.3. Social IQ and ADOS

```{r setup, include=FALSE}

View(d2)
mean(v1$VerbalIQ, na.rm = TRUE)

#make a dataframe only with visit 1 and asds

d2 = subset(v1, ASD ==1)


#remove NAs
d3 = d2[complete.cases(d2), ]


# Verbal IQ

mean(d3$VerbalIQ)
sd(d3$VerbalIQ)

m1 = map(
  alist(
  VerbalIQ~dnorm( mu, sigma ),
  mu <- a + b*ADOS,
  a ~dnorm( 18 , 10 ), #giving more room to deviate
  b ~dnorm( 0 , 5 ), #being agnostic, not really skeptic neither optimistc
  sigma~dunif( 0 , 50 )
),
data=d3 )
precis(m1)

# a person 1 ADOS point higher will fall 1,10 points in verbal IQ. 89 % of the posterior probability lies between -1.40 and -0.80. Suggesting a negative correlation
#sigma = 95 % of plausible ADOS points lies within 5,19 points of the mean.


#Nonverbal IQ

mean(d3$NonVerbalIQ)
sd(d3$NonVerbalIQ)

m2 = map(
  alist(
  NonVerbalIQ~dnorm( mu, sigma ),
  mu <- a + b*ADOS,
  a ~dnorm( 27 , 7 ), #giving more room to deivate
  b ~dnorm( 0 , 5 ), #being agnostic, not really skeptic neither optimistc
  sigma~dunif( 0 , 50 )
),
data=d3 )

precis(m2)

# A person rising in ados point will fall 0.61 point in nonverbal IQ, indicating a negative correlation w/ 89 % of the posterior probability lying between -0.86 and -0.36.
# An SD of the error is 4.35 of the mean.


#Social IQ

mean(d3$SocialIQ)
sd(d3$SocialIQ)

m3 = map(
  alist(
  SocialIQ~dnorm( mu, sigma ),
  mu <- a + b*ADOS,
  a ~dnorm( 77 , 13 ), #giving more room to deivate
  b ~dnorm( 0 , 5 ), #being agnostic, not really skeptic neither optimistc
  sigma~dunif( 0 , 50 )
),
data=d3 )

precis(m3)

# Rise in ADOS correlates negatively with social IQ, with 1 rise in ados resulting in -1.47 points in social IQ. 89% posterior probability lies between -1.95% and -0.99. And SD of the error is at 8.34.



#A model to see the correlations w/ updates priors


corm <- map(
alist(
ADOS ~ dnorm( mu , sigma ) ,
mu <- a + bV*VerbalIQ + bN*NonVerbalIQ + bS*SocialIQ ,
a ~ dnorm( 14 , 5 ) ,
bV ~ dnorm( 33 , 5 ) ,
bN ~ dnorm( 11 , 4 ) ,
bS ~ dnorm( 98 , 8 ) ,
sigma ~ dunif( 0 , 5 )
) ,
data = d3 )
precis( corm, corr = TRUE )



#plotting (see slides)

#Model 1
#plot( VerbalIQ ~ ADOS , data=d3 )
abline( a=coef(m1)["a"] , b=coef(m1)["b"] )


# display raw data and sample size

N = 30
post1 <- extract.samples( m1 , n=20 )

plot( d3$VerbalIQ , d3$ADOS ,
xlim=range(d3$ADOS) , ylim=range(d3$VerbalIQ) ,
col=rangi2 , xlab="ADOS" , ylab="VerbalIQ" )
mtext(concat("N = ",N))
# plot the lines, with transparency
for ( i in 1:20 )
abline( a=post1$a[i] , b=post1$b[i] , col=col.alpha("black",0.3) )




#Model 2
#plot( NonVerbalIQ ~ ADOS , data=d3 )
abline( a=coef(m2)["a"] , b=coef(m2)["b"] )



# display raw data and sample size

N = 30
post2 <- extract.samples( m2 , n=20 )

plot( d3$NonVerbalIQ , d3$ADOS ,
xlim=range(d3$ADOS) , ylim=range(d3$NonVerbalIQ) ,
col=rangi2 , xlab="ADOS" , ylab="NonVerbalIQ" )
mtext(concat("N = ",N))
# plot the lines, with transparency
for ( i in 1:20 )
abline( a=post2$a[i] , b=post2$b[i] , col=col.alpha("black",0.3) )





#Model 3
#plot( SocialIQ ~ ADOS , data=d3 )
abline( a=coef(m3)["a"] , b=coef(m3)["b"] )

# display raw data and sample size

N= 30
post3 <- extract.samples( m3 , n=20 )

plot( d3$SocialIQ , d3$ADOS ,
xlim=range(d3$ADOS) , ylim=range(d3$SocialIQ) ,
col=rangi2 , xlab="ADOS" , ylab="SocialIQ" )
mtext(concat("N = ",N))
# plot the lines, with transparency
for ( i in 1:20 )
abline( a=post3$a[i] , b=post3$b[i] , col=col.alpha("black",0.3) )


grid.arrange(post1, post2, post3)

gridExtra::grid.arrange(post1,post2,post3)
```

2. Do the different aspects of IQ account for different portions of the variance in ADOS? 
2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?
2.2. Build the model, assess its quality, write a few lines interpreting the results.

```{r setup, include=FALSE}
#Do the different aspects of IQ account for different portions of the variance in ADOS?

#well if two of the betas are very correlated, then they explain the same phenomena and we cannot use it

# y = a + bV(erbaliq) + bN(onVerbaliq) + bS(ocialIQ)


#Does it make sense to have all IQ measures in the same model? Motivate answer and do it.

#Fit Model

#scaling 

scale(d3$VerbalIQ)
scale(d3$NonVerbalIQ)
scale(d3$SocialIQ)

d4 = scale(d3)
mean(d3$VerbalIQ)
sd(d3$VerbalIQ)


m4 <- map(
alist(
VerbalIQ ~ dnorm( mu , sigma ) ,
mu <- a + bN*NonVerbalIQ  ,
a ~ dnorm( 18 , 8 ) ,
bN ~ dnorm( 0 , 5 ) ,
sigma ~ dunif( 0 , 50 )
) ,
data = d3 )
precis(m4, corr = TRUE)

#### 0.69


mean(d3$NonVerbalIQ)
sd(d3$NonVerbalIQ)

m5 <- map(
alist(
NonVerbalIQ ~ dnorm( mu , sigma ) ,
mu <- a + bS*SocialIQ  ,
a ~ dnorm( 27 , 5 ) ,
bS ~ dnorm( 0 , 5 ) ,
sigma ~ dunif( 0 , 50 )
) ,
data = d3 )
precis(m5, corr = TRUE)

#### 0.07

mean(d3$SocialIQ)
sd(d3$SocialIQ)

m6 <- map(
alist(
SocialIQ ~ dnorm( mu , sigma ) ,
mu <- a + bV*VerbalIQ  ,
a ~ dnorm( 78 , 5 ) ,
bV ~ dnorm( 0 , 5 ) ,
sigma ~ dunif( 0 , 50 )
) ,
data = d3 )
precis(m6, corr = TRUE)

#### 0.60


#### The betas are not completely correlated, suggesting the G factor doe snot explain all. Nonverbal + social IQ are very little correlated meaning there is very little shared variance. Those would go good in a model together, since they express two difference phenomenas.


m7 <- map(
alist(
ADOS ~ dnorm( mu , sigma ) ,
mu <- a + bN*NonVerbalIQ + bS*SocialIQ ,
a ~ dnorm( 0 , 5 ) ,
bN ~ dnorm( 0 , 1 ) , #lets try to be more liberal
bS ~ dnorm( 0, 1 ) , 
sigma ~ dunif( 0 , 50 )
) , 
data = as.data.frame(d4) ) 
precis(m7, corr = TRUE)


plot(precis(m7))
dens(precis(m7))

#Quality assement
#Make density plots from sim values

sim.Ados <- sim(m7, data = as.data.frame(d4))
#)
dens(sim.Ados, col = "red", xlim = c(-2, 2.5), ylim = c(0, 1),  xlab = "ADOS")
par(new=TRUE)
dens(as.data.frame(d4)$ADOS, xlim = c(-2,2.5), ylim = c(0,1), xlab = "ADOS")


#Let's quality check a model with all three predictors

m8 <- map(
alist(
ADOS ~ dnorm( mu , sigma ) ,
mu <- a + bN*NonVerbalIQ + bS*SocialIQ + bV*VerbalIQ,
a ~ dnorm( 0 , 5 ) ,
bN ~ dnorm( 0 , 1 ) , #lets try to be more liberal
bS ~ dnorm( 0, 1 ) ,
bV ~ dnorm( 0, 1 ) ,
sigma ~ dunif( 0 , 50 )
) , 
data = as.data.frame(d4) ) 
precis(m8, corr = TRUE)


plot(precis(m8))

sim.Ados <- sim(m8, data = as.data.frame(d4))
#)
dens(sim.Ados, col = "red", xlim = c(-2, 2.5), ylim = c(0, 1),  xlab = "ADOS")
par(new=TRUE)
dens(as.data.frame(d4)$ADOS, xlim = c(-2,2.5), ylim = c(0,1), xlab = "ADOS")


#89% interval = the probability that the true result given the model 

```


3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.

```{r setup, include=FALSE}

all = v1[complete.cases(v1), ]

#all scaled (centered)
scall = scale(all)

View(scall)

#interaction model
m9 <- map(
  alist(
    ADOS ~ dnorm( mu , sigma ) ,
    mu <- a + bN*NonVerbalIQ + bS*SocialIQ +    bNS*NonVerbalIQ*SocialIQ ,
    a ~ dnorm( 0 , 5 ) ,
    bN ~ dnorm( 0 , 1 ) , #lets try to be more liberal
    bS ~ dnorm( 0, 1 ) ,
    bNS ~ dnorm( 0, 1 ) ,
    sigma ~ dunif( 0 , 50 )
  ) , 
data= as.data.frame(scall))

precis(m9)


#of the two models above withut interactions m7 is the ost appropriate


#Model without interaction
m7.2 <- map(
alist(
ADOS ~ dnorm( mu , sigma ) ,
mu <- a + bN*NonVerbalIQ + bS*SocialIQ ,
a ~ dnorm( 0 , 5 ) ,
bN ~ dnorm( 0 , 1 ) , #lets try to be more liberal
bS ~ dnorm( 0, 1 ) , 
sigma ~ dunif( 0 , 50 )
) , 
data = as.data.frame(scall) ) 




compare (m7.2, m9)

(ados.models = compare(m7.2,m9))
plot(ados.models, SE=TRUE, dSE=TRUE)

```


4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?
4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.

```{r setup, include=FALSE}


```

5. Bonus questions: Including measurement errors. 
5.1. Let's assume that ADOS has a measurement error of 1. How would you model that in any of the previous models? 
5.2. We know that IQ has an estimated measurement error of 2.12. How would you include that? 


